---
title: "The Effect of Vitamin C on Tooth Growth in Guinea Pigs"
author: "Matthew Bender"
date: "August 19th, 2015"
output: html_document
---

### Overview
The purpose of the this data analysis is to analyze the ToothGrowth data set by
comparing the guinea tooth growth by supplement and dose. First, I will do
exploratory data analysis on the data set. Then I will do the comparison with
confidence intervals in order to make conclusions about the tooth growth.

### Load the ToothGrowth data and perform exploratory data analyses
```{r results='hide'}
library(datasets)
data(ToothGrowth)
str(ToothGrowth)
head(ToothGrowth)
summary(ToothGrowth)
```

```{r}
library(ggplot2)
t = ToothGrowth
levels(t$supp) <- c("Orange Juice", "Ascorbic Acid")
ggplot(t, aes(x=factor(dose), y=len)) + 
  facet_grid(.~supp) +
  geom_boxplot(aes(fill = supp), show_guide = FALSE) +
  labs(title="Guinea pig tooth length by dosage for each type of supplement", 
    x="Dose (mg/day)",
    y="Tooth Length")
```

### Basic summary of the data
The box plots seem to show, increasing the dosage increases the tooth growth. Orange
juice is more effective than ascorbic acid for tooth growth when the dosage is .5
to 1.0 milligrams per day. Both types of supplements are equally as effective 
when the dosage is 2.0 milligrams per day.

### Use confidence intervals & hypothesis tests to compare tooth growth by supplement and dose
#### Hypothesis #1
Orange juice & ascorbic acid deliver the same tooth growth across the data set.
```{r}
hypoth1<-t.test(len ~ supp, data = t)
hypoth1$conf.int
hypoth1$p.value
```
The confidence intervals includes 0 and the p-value is greater than the 
threshold of 0.05.  The null hypothesis cannot be rejected.

#### Hypothesis #2
For the dosage of 0.5 mg/day, the two supplements deliver the same tooth growth.
```{r}
hypoth2<-t.test(len ~ supp, data = subset(t, dose == 0.5))
hypoth2$conf.int
hypoth2$p.value
```
The confidence interval does not include 0 and the p-value is below the 0.05 
threshold. The null hypothesis can be rejected. The alternative hypothesis 
that 0.5 mg/day dosage of orange juice delivers more tooth growth than ascorbic 
acid is accepted.

#### Hypothesis #3
For the dosage of 1 mg/day, the two supplements deliver the same tooth growth
```{r}
hypoth3<-t.test(len ~ supp, data = subset(t, dose == 1))
hypoth3$conf.int
hypoth3$p.value
```
The confidence interval does not include 0 and the p-value is smaller than the 
0.05 threshold. The null hypothesis can be rejected. The alternative hypothesis 
that 1 mg/day dosage of orange juice delivers more tooth growth than ascorbic 
acid is accepted.

#### Hypothesis #4
For the dosage of 2 mg/day, the two supplements deliver the same tooth growth
```{r}
hypoth4<-t.test(len ~ supp, data = subset(t, dose == 2))
hypoth4$conf.int
hypoth4$p.value
```
The confidence interval does include 0 and the p-value is larger than the 
0.05 threshold. The null hypothesis cannot be rejected.


### Conclusions & assumptions
Orange juice delivers more tooth growth than ascorbic acid for dosages 0.5 & 
1.0. Orange juice and ascorbic acid deliver the same amount of tooth growth for 
dose amount 2.0 mg/day.  For the entire data set we cannot conclude orange juice
is more effective that ascorbic acid.

Assumptions

* Normal distribution of the tooth lengths
* No other unmeasured factors are affecting tooth length